---
title: "HW_6"
author: "Amy Liu"
date: "2023-11-29"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(p8105.datasets)
library(modelr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1
* Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r}
homicide = homicide_df |>
  mutate(city_state = paste0(city,", ",state),
         solved = ifelse(disposition == "Closed by arrest",1,0)) |>
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO",
                            "Tulsa, AL")) |>
  filter(victim_race %in% c("White", "Black")) |>
  mutate(victim_age = as.numeric(victim_age))


homicide_df 
```

*For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
baltimore = homicide |>
  filter(city == "Baltimore")

model_baltimore = glm(solved ~ victim_age + victim_race + victim_sex,
                      data = baltimore, family = binomial())

model <- model_baltimore |>
  broom::tidy() |>
  knitr::kable(digits = 3)
model
save(model, file = "./Baltimore Logistic Regression.RData")


```

